Quantum Pseudo-Telepathy

### 2023-06-04 paper/main.tex
quantum initial state + grid

@@

What if Alice and Bob decide for a probabilistic strategy? That is,
Alice and Bob each carry a coin. When assigned a row/column, they flip
their coins and fill the entries of the row/column based on the
outcomes of the coins. Can they do better? To randomly assign values
to one row (equivalently, one column) is essentially the same as
randomly selecting one of the \(2^{9}\) possible predefined
\(3 \times 3\) grids (including those matrices that would lead to
invalid answers). In other words, a probabilistic strategy for the
Magic Square game is one in which Alice and Bob randomly select a
deterministic strategy. Hence, no matter how lucky Alice and Bob are,
the probability of success of any strategy that they come up with is
bounded by the winning probability of the best deterministic
strategy, which is
\(\nicefrac{8}{9}\). Theorem~\ref{thm:classical-success} formalizes
this argument.
\subsection{Quantum Solution}
\label{sec:quantum-sol}

>>>>

\usepackage{tikz}
\usetikzlibrary{quantikz2}

\newcommand{\ketbra}[2]{\ensuremath{| #1 \rangle \langle #2 |}}
\newcommand{\ci}{\mathrm{i}}
\DeclarePairedDelimiter{\card}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\DeclareMathOperator{\tr}{tr}

@@

What if Alice and Bob decide for a probabilistic strategy? That is,
Alice and Bob each carry a coin. When assigned a row/column, they flip
their coins and fill the entries of the row/column based on the
outcomes of the coins. Can they do better? To randomly assign values
to one row (equivalently, one column) is essentially the same as
randomly selecting one of the \(2^{9}\) possible predefined
\(3 \times 3\) grids (including those matrices that would lead to
invalid answers). In other words, a probabilistic strategy for the
Magic Square game is one in which Alice and Bob randomly select a
deterministic strategy. Hence, no matter how lucky Alice and Bob are,
the probability of success of any strategy that they come up with is
bounded by the winning probability of the best deterministic
strategy, which is
\(\nicefrac{8}{9}\). Theorem~\ref{thm:classical-success} formalizes
this argument.

\subsection{Quantum Solution}
\label{sec:quantum-sol}
The quantum strategy for the Magic Square game consists of Alice and
Bob carrying entangled qubits with them. For each round, they need a
pair of qubits each. Then, when Charlie asks Alice (respectively Bob)
to fill the entries of the \(i\)-th row (respectively \(j\)-th
column), Alice measures her two qubits three times. The outcome of
each measurement gives the value for each cell (in practice, they do
not even need the third measurement, since they can determine the third
value from the first two). Key to this strategy is that the outcomes
of all possible measurements are \(\pm 1\). The initial, entangled state
is (subscripts \(A\) and \(B\) indicate, respectively,
Alice's and Bob's qubits)
\begin{equation}
  \ket{\psi} = \frac{1}{2}\left(
    \ket{00}_{A} \otimes \ket{00}_{B} + \ket{01}_{A} \otimes \ket{01}_{B} +
    \ket{10}_{A} \otimes \ket{10}_{B} + \ket{11}_{A} \otimes \ket{11}_{B}
  \right).
\end{equation}
Figure~\ref{fig:circuit-psi} shows the quantum circuit that can be
used to prepare the initial state \(\ket{\psi}\).

\begin{figure}[htbp]
  \centering
  \begin{quantikz}
    \lstick{\ket{0}} & \gate{H}
    \slice[label style={xshift=-15mm}]
    {\(\frac{1}{2}(\ket{00} + \ket{01} + \ket{10} +
      \ket{11}) \otimes \ket{00}\)}
    &&&&& \ctrl{2} &&
    \\
    \lstick{\ket{0}} & \gate{H} &&&&&& \ctrl{2} & \\
    \lstick{\ket{0}} & &&&&& \targ{} && \\
    \lstick{\ket{0}} & &&&&&& \targ{}
    \slice[label style={xshift=5mm}]
    {\(\frac{1}{2}(\ket{0000} + \ket{0101} + \ket{1010} +
      \ket{1111})\)}
    &
  \end{quantikz}
  \caption{\label{fig:circuit-psi}}
\end{figure}

As seen in Section~\ref{sec:classic-sol}, the \emph{magic} square is
called this way because, under classical assumptions, there is no such
a square: it is impossible to construct a \(3 \times 3\) matrix filled
with \(\pm 1\)'s, with the constraints that the product of every row
is \(+1\) and the product of every column is \(-1\). We can, however,
construct a ``quantum magic'' square. Figure~\ref{fig:quantum} shows
the \(3 \times 3\) grid that Alice and Bob can use to win the Magic
Square game systematically.  Notice that the eigenvalues of the
observables at each cell are \(\pm 1\). Furthermore, the product of
every row is \(+I\) and the product of every column is \(-I\). (The
original observables did not have the negative \(X\) and \(Z\) in the
last row, so these two properties are not mandatory~\cite{Mermin1990,
  aravind:xxx}.  We have adapted the observables merely to be
``consistent'' with the properties of the classical magic square, but
what really matters for Alice and Bob are the outcomes of the
measurements, which they will use to fill the entries asked by
Charlie.)

{\renewcommand{\arraystretch}{2}
\begin{figure}
  \centering
  \[\begin{array}{|*{3}{>{\quad}c<{\quad}|}>{\quad}l<{\quad}}
      \hhline{---~}
      +I \otimes Z & +Z \otimes I & +Z \otimes Z & \prod = I \otimes I = I
      \\ \hhline{---~}
      +X \otimes I & +I \otimes X & +X \otimes X & \prod = I \otimes I = I
      \\ \hhline{---~}
      -X \otimes Z & -Z \otimes X & +Y \otimes Y & \prod = -\ci{}I \otimes \ci{}I = I
      \\ \hhline{---~}
      \multicolumn{1}{>{\quad}c<{\quad}}{\prod = -I \otimes I = -I}
      & \multicolumn{1}{>{\quad}c<{\quad}}{\prod = -I \otimes I = -I}
      & \multicolumn{1}{>{\quad}c<{\quad}}{\prod = \ci{}I \otimes \ci{}I = -I}
    \end{array}\]
  \caption{\label{fig:quantum}}
\end{figure}}
